Parameter estimation

### 2023-06-20 SFE-article/Sections/Results.tex
20.06

    \begin{itemize}
        \item Partition coefficient:\qquad\quad\qquad$k_m$
        \item Internal Diffusion coefficient: \quad$D_i$
        \item Axial Diffusion coefficient: \qquad$D_e^M$
        \item Saturation concentration: \quad\qquad$C_{sat}$
        \item Standard Deviation: \qquad\qquad\quad$\sigma$
    \end{itemize}

@@

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lccccccc }
			\hline 
			Parameter		&$k_m$[-] 	& $D_i$[m/s$^2$] 	& $D_e^M$[m/s$^2$]	& $C_{sat}$ [kg/m$^3$] 	& $m_{total}$[g]	& $\tau$[-] 	& $\sigma$[-] \\  \hline
			Lower bound		&0	  		& 0 	  			& 0 				&	0					& 80 		 		& 0 	   		& 0 \\ 
			Upper bound		&1 			& $+\infty$ 		& $+\infty$			&	$+\infty$			& 150 				& 1 			& $+\infty$ \\ 
			Initial guess	&1	  		& 3 	  			& 1 				&	3					& 80 		 		& 0.65 	   		& 0.1 \\  \hline
		\end{tabular} }
	\caption{Constraints and initial guess of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

@@

	The further spifflication of the process model can be introduced based on $C_{sat}$ results. Values of saturation concentrations are at the level of 1e5 which is an internal upper bound of the optimizer used if $+\infty$ indicates the upper bound. This suggest that solvent hasn't reach the saturation concentration and the process model can be simplified by replacing $\gamma(c_f)$ function with unity.

@@

	The results obtained from the parameter estimation are used to develop a generalized process model. The model generalization is performed by introducing correlations which link the estimated parameters and the operating conditions. Since the experiments were performed under varying temperature and pressure, the influence of these variables is taken into account. To simplify the problem, the correlations are written as functions of fluid density rather than temperature and pressure. Since the optimizer found that $k_m$, $m_{total}$ and $C_{sat}$ don't change with the density variations, they are assumed to be constants and not considered further.

>>>>

    \begin{itemize}
        \item Partition coefficient:\qquad\quad\qquad$k_m$
        \item Internal diffusion coefficient: \quad$D_i$
        \item Axial diffusion coefficient: \qquad$D_e^M$
        \item Limiting concentration: \quad\qquad$C_{lim}$
        \item Standard deviation: \qquad\qquad\quad$\sigma$
    \end{itemize}

@@

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lccccccc }
			\hline 
			Parameter		&$k_m$[-] 	& $D_i$[m/s$^2$] 	& $D_e^M$[m/s$^2$]	& $C_{lim}$ [kg/m$^3$] 	& $m_{total}$[g]	& $\tau$[-] 	& $\sigma$[-] \\  \hline
			Lower bound		&0	  		& 0 	  			& 0 				&	0					& 80 		 		& 0 	   		& 0 \\ 
			Upper bound		&1 			& $+\infty$ 		& $+\infty$			&	$+\infty$			& 150 				& 1 			& $+\infty$ \\ 
			Initial guess	&1	  		& 3 	  			& 1 				&	3					& 80 		 		& 0.65 	   		& 0.1 \\  \hline
		\end{tabular} }
	\caption{Constraints and initial guess of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

@@

	The further spifflication of the process model can be introduced based on $C_{lim}$ results. Values of saturation concentrations are at the level of 1e5 which is an internal upper bound of the optimizer used if $+\infty$ indicates the upper bound. This suggest that solvent hasn't reach the saturation concentration and the process model can be simplified by replacing $\gamma(c_f)$ function with unity.

@@

	The results obtained from the parameter estimation are used to develop a generalized process model. The model generalization is performed by introducing correlations which link the estimated parameters and the operating conditions. Since the experiments were performed under varying temperature and pressure, the influence of these variables is taken into account. To simplify the problem, the correlations are written as functions of fluid density rather than temperature and pressure. Since the optimizer found that $k_m$, $m_{total}$ and $C_{lim}$ don't change with the density variations, they are assumed to be constants and not considered further.

### 2023-06-20 SFE-article/Sections/Saturation.tex
20.06

	\subsubsection{Saturation Concentration}

	The above model does not take into account the saturation of fluid, which can be introduced by multiplying the gradient by a function $\gamma(c_f)$ (equation \ref{EQ: C_sat_function}). $\gamma(c_f)$ describe the reverse logistic function, which is equal to unity below the $c_{sat}$, the saturation concentration, and equal to zero, above the $c_{sat}$. The $\gamma(c_f)$ for $c_{sat}=5$ is shown on figure \ref{fig: Gamma_function}.

	{\footnotesize
		\begin{equation}
			\gamma(c_f) = \cfrac{1}{1+\exp ( -k_{sat} \left( c_f - c_{sat} \right) ) }
			\label{EQ: C_sat_function}
			\end{equation}
	}

	where $k_{sat}$ is the growth rate, and it is defined as $k_{sat} = 4$. The growth rate corresponds for the shape of the $\gamma(c_f)$ function. The bigger value of the $k_{sat}$ the steeper the function is. In the limit when $ k_{sat}\rightarrow+\infty$ become a step function. If the growth rate is to high, the gradient of the $\gamma$ function become high, or goes to infinity, which might cause difficulties for a gradient-based optimizer.

@@

				\begin{axis}[
					legend style={draw=none},
					xmin = 0, xmax = 10,
					ymin = -0.5, ymax = 1.5,
					xlabel = {$c_f[kg/m^3]$},
					ylabel = {$\gamma(c_f)[-]$}]
					\addplot[
					domain = 0:10,
					samples = 100,
					smooth,
					thick,
					blue,
					] { 1 / (1 + exp( 4 .* ( x - 5 ) ) ) };
					\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(5,-0.5)(5,1.5)};

					\addlegendentry{$\gamma(c_f)$}
					\addlegendentry{$c_{sat}$}
				\end{axis}

			\end{tikzpicture}
			\caption{$\gamma(c_f)$ function under assumption of $c_{sat}=5 [kg/m^3]$}
			\label{fig: Gamma_function}
	\end{figure}

@@

	{\footnotesize
		\begin{equation}
			\label{Model_kinetic}
				{\color{blue}r_e}(t,z) = -\cfrac{{\color{orange}D_i}({\color{blue}T}(t,z), {\color{red}P}(t))}{{\color{magenta} \mu l^2} } \gamma(c_f) \left({\color{blue}c_s}(t,z) - \cfrac{{\color{magenta}\rho_s}}{{\color{orange}k_m}({\color{blue}T}(t,z)){\color{orange}\rho}({\color{blue}T}(t,z),{\color{red}P}(t))}  {\color{blue}c_f}(t,z) \right)
		\end{equation} }

>>>>

	\subsubsection{Concentration limit}

	To mimic that not all solute might be obtained at specific operating conditions, the coefficient $\gamma$ is introduced to limit the available oil. The $\gamma(c_s)$ describe the reverse logistic function, which is equal to unity above the $c_{lim}$, the saturation concentration, and equal to zero, below the $c_{lim}$. The $\gamma$ parameter multiply the solute concentration term in solid phase in the concentration gradient. As the concentration of solute decrease, it can reach the limiting value $c_{lim}$ when the $\gamma$ goes to zero and stop the extraction from the solid.
	The $\gamma(c_f)$ for $c_{sat}=5$ is shown on figure \ref{fig: Gamma_function}.

	{\footnotesize
		\begin{equation}
			\gamma(c_s) = \cfrac{1}{1+\exp ( -k_{sat} \left( c_s - c_{lim} \right) ) }
			\label{EQ: C_sat_function}
			\end{equation}
	}

	where $k_{lim}$ describe the curvature of the $\gamma$ function, and it is defined as $k_{lim} = 1$. The bigger value of the $k_{sat}$ the steeper the function is. In the limit when $ k_{lim}\rightarrow+\infty$ become a step function. If the growth rate is to high, the gradient of the $\gamma$ function become high, or goes to infinity, which might cause difficulties for a gradient-based optimizer.

@@

				\begin{axis}[
					legend style={draw=none},
					xmin = 0, xmax = 100,
					ymin = -0.5, ymax = 1.5,
					xlabel = {$c_f[kg/m^3]$},
					ylabel = {$\gamma(c_f)[-]$}]
					\addplot[
					domain = 0:100,
					samples = 100,
					smooth,
					thick,
					blue,
					] { 1 / (1 + exp( 1 * ( 50 - x ) ) ) };
					\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(50,-0.5)(50,1.5)};

					\addlegendentry{$\gamma(c_s)$}
					\addlegendentry{$c_{lim}$}
				\end{axis}

			\end{tikzpicture}
			\caption{$\gamma(c_s)$ function under assumption of $c_{lim}=50~[kg/m^3]$}
			\label{fig: Gamma_function}
	\end{figure}

@@

	{\footnotesize
		\begin{equation}
			\label{Model_kinetic}
				{\color{blue}r_e}(t,z) = -\cfrac{{\color{orange}D_i}({\color{blue}T}(t,z), {\color{red}P}(t))}{{\color{magenta} \mu l^2} } \left({\color{blue}c_s}(t,z) \gamma(c_s) - \cfrac{{\color{magenta}\rho_s}}{{\color{orange}k_m}({\color{blue}T}(t,z)){\color{orange}\rho}({\color{blue}T}(t,z),{\color{red}P}(t))}  {\color{blue}c_f}(t,z) \right)
		\end{equation} }

### 2023-06-20 SFE-article/Sections/Thermo_imp.tex
20.06

	Supercritical $CO_2$'s thermodynamic properties, such as density, the local speed of sound, and specific heat capacity, vary significantly with slight changes in temperature and pressure due to real gas effects. The Peng-Robinson equation of state (P-R EOS) is used to calculate the thermodynamic properties by accounting for these real gas effects. The P-R EOS belongs to a specific class of thermodynamic models for modelling the pressure of a gas as a function of temperature and density. It can be written as a cubic function of the molar volume (of the density). Detail information about Peng-Robinson equation of state can be found in the work of \citet{Peng1976}, \citet{Elliott2011} or \citet{Pratt2001}. The P-R EOS is presented by Equation \ref{EQ: PR}

	{\footnotesize
		\begin{equation}
			P = \cfrac{RT}{V_m - b} - \cfrac{a\alpha}{V_m^2 + 2bV_m - b^2}
			\label{EQ: PR}
		\end{equation}
	}

	The parameters $a, b, \alpha$ are parameters defined as presented in the Appendix \ref{subsubsec: Equation of state}.

	The properties of $CO_2$  are presented as a function of operating conditions (temperature and pressure) in Figure \ref{fig: SFE_Properties}. At standard atmospheric pressure and temperature, $CO_2$  behaves as an ideal gas, and its compressibility factor equals unity. However, at high pressures and/or low temperatures, intermolecular forces between gas molecules become more significant, causing them to deviate from ideal behaviour. As a result, the compressibility factor can either be greater than or less than unity, depending on the magnitude of these forces. As presented in Figure \ref{fig: SFE_Properties_Compressibility}, the compressibility factor obtained from the Peng-Robinson equation of state varies strongly depending on the operating conditions. The compressibility factor can be obtained by solving the polynomial form of the P-R EOS given by Equation \ref{EQ: PR_polynomial}.

	{\footnotesize
		\begin{equation}
			Z^3 - (1-B)Z^2+(A-2B-3B^2)Z -(AB-B^2-B^3) = 0
			\label{EQ: PR_polynomial}
		\end{equation}
	}

	where $A$ and $B$ are parameters as defined in the Appendix \ref{subsubsec: Equation of state}. The roots of the polynomial can be found iteratively or by the Cardano formula. In a one-phase region, the fluid is described by one real root corresponding to the gas, liquid or supercritical phase. The gas-liquid mixture is present in the two-phase region, and two roots are found. The biggest root is assigned to the gas phase, and the smallest root corresponds to the liquid phase.

	The real gas effects are also visible on the density plot presented in Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor and its value depends on the operating conditions. The fluid's properties near the critical point are unique and combine gas-like and liquid-like properties. The details of calculations are explained in the Appendix \ref{subsubsec: Fluid density}.

@@

	To address this issue, the correlations of transport coefficients are either empirical or based on some theoretical foundation. Chapman-Enskog's theory (presented in \citet{Chapman1991}) for transport properties of dense gases based on the distribution function is a popular theoretical approach. However, the Chapman-Enskog theory was developed for rigid spherical molecules and modifications are required to apply it to real gases. Many correlations have been proposed following the Chapman-Enskog theory in the form of reduced density and reduced temperature, such as those developed by \citet{Fenghour1998} and \citet{Laesecke2017} from the National Institute of Standards and Technology (NIST). A comparison of these correlations is presented in Figure \ref{fig: SFE_Properties_Viscosity}.

@@

	Similarly, several correlations for thermal conductivity of $CO_2$ were compared in Figure \ref{fig: SFE_Properties_Thermal}. The presented figures show regions around the critical point where the singularity is present. Similarities between specific heat and thermal conductivity can be observed. The NIST correlation (\citet{Huber2016}) captures the singular behaviour of thermal conductivity around the critical point. The correlations are applicable for the temperature range from the triple point to 1100 $K$ and pressures up to 200 $MPa$.

>>>>

	Supercritical $CO_2$'s thermodynamic properties, such as density, the local speed of sound, and specific heat capacity, vary significantly with slight changes in temperature and pressure due to real gas effects. The Peng-Robinson equation of state (P-R EOS) is used to calculate the thermodynamic properties by accounting for these real gas effects are presented Appendix \ref{subsubsec: Equation of state}. The P-R EOS belongs to a specific class of thermodynamic models for modelling the pressure of a gas as a function of temperature and density.

	The properties of $CO_2$  are presented as a function of operating conditions (temperature and pressure) in Figure \ref{fig: SFE_Properties}. At standard atmospheric pressure and temperature, $CO_2$  behaves as an ideal gas, and its compressibility factor equals unity. However, at high pressures and/or low temperatures, intermolecular forces between gas molecules become more significant, causing them to deviate from ideal behaviour. As a result, the compressibility factor can either be greater than or less than unity, depending on the magnitude of these forces. As presented in Figure \ref{fig: SFE_Properties_Compressibility}, the compressibility factor obtained from the Peng-Robinson equation of state varies strongly depending on the operating conditions.

	The real gas effects are also visible on the density plot presented in Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor and its value depends on the operating conditions. The fluid properties near the critical point are unique and combine gas-like and liquid-like properties. The details of calculations are explained in the Appendix \ref{subsubsec: Fluid density}.

@@

	To address this issue, the correlations of transport coefficients are either empirical or based on some theoretical foundation. Chapman-Enskog's theory (presented in \citet{Chapman1991}) for transport properties of dense gases based on the distribution function is a popular theoretical approach. However, the Chapman-Enskog theory was developed for rigid spherical molecules and modifications are required to apply it to real gases. Many correlations have been proposed following the Chapman-Enskog theory in the form of reduced density and reduced temperature, such as those developed by \citet{Fenghour1998}, and \citet{Laesecke2017} from the National Institute of Standards and Technology (NIST). The correlation of \citet{Laesecke2017} is presented in Figure \ref{fig: SFE_Properties_Viscosity}.

@@

%	Similarly, several correlations for thermal conductivity of $CO_2$ were compared in Figure \ref{fig: SFE_Properties_Thermal}. The presented figures show regions around the critical point where the singularity is present. Similarities between specific heat and thermal conductivity can be observed. The NIST correlation (\citet{Huber2016}) captures the singular behaviour of thermal conductivity around the critical point. The correlations are applicable for the temperature range from the triple point to 1100 $K$ and pressures up to 200 $MPa$.

	Similarly, the NIST developed the correlation, which describe the behaviour of thermal diffusivity of $CO_2$. The work of \citet{Huber2016} captures the singular behaviour of thermal conductivity around the critical point. The correlation is applicable for the temperature range from the triple point to 1100 $K$ and pressures up to 200 $MPa$. Figure \ref{fig: SFE_Properties_Thermal}. The presented figure show regions around the critical point where the singularity is present.

### 2023-06-20 SFE-article/Sections/Thermodynamic_details.tex
20.06

	\subsubsection{Equation of state and properties of the fluid phase} \label{subsubsec: Equation of state}
	In this study, the influence of real gas effects are introduced through ${\color{red}P}(t) {\color{orange}v_m}[T(t,z),P(t)] = {\color{orange}Z}{\color{magenta}R}{\color{blue}T}(t,z)$, where ${\color{orange}v_m}$ represents the molar volume of $\text{CO}_2$, ${\color{orange}Z}$ denotes its compressibility factor, and $R$ is the universal gas constant. The main advantage of using compressibility in calculations is to express it as an explicit function of temperature and pressure. This is possible when $Z$ is obtained as one of the physically meaningful roots of a polynomial equation, such as the Van der Waals, Peng-Robinson, or Redlich-Kwong equation of state. In this study, the Peng-Robinson equation of state (\citet{Peng1976}) was selected.

	In the Peng-Robinson equation of state, the compressibility ${\color{orange}\overline{Z}}[{\color{blue}T}(t,z), {\color{red}P}(t)]$ solves the third-order polynomial equation shown in Equation \ref{eq: Compressibility}. Numerical methods such as Newton-Raphson can be used to solve the polynomial equation, but the Cardano formula is preferred as it yields an analytical solution. Further details regarding the Cardano formula can be found in Appendix \ref{CH: Cardano}.

@@

	{\footnotesize
	\begin{subequations} \label{eq: PR_AB}
		\begin{align} 
			{\color{orange}A}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]	& = \dfrac{{\color{orange}\alpha}\left[T\left(t,z\right)\right]{\color{magenta}a^c}{\color{red}P}(t)}{{\color{magenta}R}^2{\color{blue}T}^2(t,z)};													\label{eq: PR_A}\\
			{\color{orange}B}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]	& = \dfrac{{\color{magenta}b}{\color{red}P}(t)}{{\color{magenta}R}{\color{blue}T}(t,z)}	\label{eq: PR_B}.
		\end{align}
	\end{subequations}}

	The quantity ${\color{orange}\alpha}\left[{\color{blue}T}\left(t,z\right)\right]$ is a dimensionless correction term that depends on the temperature ${\color{blue}T}\left(t,z\right)$ and a constant ${\color{magenta}\kappa}$. The constant ${\color{magenta}\kappa}$ is given by the formula ${\color{magenta}\kappa} = 0.37464 + 1.54226 {\color{magenta}\omega} - 0.26992 {\color{magenta}\omega^2}$, where ${\color{magenta}\omega}$ is the acentric factor of $\text{CO}_2$ molecules and is equal to 0.239.

@@

@@

	where ${\color{magenta}M_{CO_2}}$ denotes the molar mass of $CO_2$, and ${\color{orange}\overline{Z}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ is the compressibility factor that solves Eq. \eqref{eq: Compressibility}. This means that the density of the fluid depends on both space and time, due to its dependence on temperature and pressure.
%	, and it can be expressed in terms of a set of physical constants, denoted by $\varphi{\rho_\text{F}} = \left({\color{magenta}R}, {\color{magenta}M_{CO_2}}, {\color{magenta}T^c{CO_2}}, {\color{magenta}P^c{CO_2}}, {\color{magenta}\omega{\text{CO}_2}}\right)$.

@@

	The specific heat $C_p^{\text{F}}$ can be calculated from the equation of state, again under the assumption that the fluid phase consists of pure carbon dioxide and that the specific heat of real fluids can be calculated from an ideal contribution plus a residual term \citet{Pratt2001}. The main steps in the derivation of ${\color{orange}C^{\text{CO}_2}_p}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ are presented below:

	{\footnotesize
	\begin{subequations}\begin{align}
			{\color{orange}C^{\text{CO}_2}_v}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right] & = {\color{orange}C_v^{\text{I}}}\left[{\color{blue}T}(t,z),P(t)\right] + {\color{orange}C_v^{\text{R}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right] \label{eq: Cv_CO2}; \\ 
			{\color{orange}C^{\text{CO}_2}_p}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right] & = \underbrace{{\color{orange}C_p^{\text{I}}}\left[{\color{blue}T}(t,z),P(t)\right]}_{\text{Eq. } \eqref{eq: Cp_I}} + \underbrace{{\color{orange}C_p^{\text{R}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]}_{\text{Eq. } \eqref{eq: Cp_R}}  \label{eq: Cp_CO2}.
	\end{align}\end{subequations} }

	${\color{orange}C^{\text{CO}_2}_v}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ and ${\color{orange}C^{\text{CO}_2}_p}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ are the specific heat of $\text{CO}_2$ at constant volume and pressure, respectively. ${\color{orange}C_v^{\text{I}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ and ${\color{orange}C_p^{\text{I}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$, with ${\color{orange}C_p^{\text{I}}}({\color{blue}T}(t,z)) - {\color{orange}C_v^{\text{I}}}({\color{blue}T}(t,z)) = {\color{magenta}R}$, are the specific heat of an ideal gas at constant volume and pressure. ${\color{orange}C_v^{\text{R}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ and ${\color{orange}C_p^{\text{R}}}({\color{blue}T}(t,z), {\color{red}P}(t))$ are the correction terms. \\

>>>>

	\subsubsection{Equation of state and properties of the fluid phase} \label{subsubsec: Equation of state}
	In this study, the influence of real gas effects are introduced through ${\color{red}P}(t) {\color{orange}v_m}({\color{blue}T}(t,z),{\color{red}P}(t)) = {\color{orange}Z}{\color{magenta}R}{\color{blue}T}(t,z)$, where ${\color{orange}v_m}$ represents the molar volume of $CO_2$, ${\color{orange}Z}$ denotes its compressibility factor, and $R$ is the universal gas constant. The main advantage of using compressibility in calculations is to express it as an explicit function of temperature and pressure. It can be written as a cubic function of the molar volume (of the density). Detail information about Peng-Robinson equation of state can be found in the work of \citet{Peng1976}, \citet{Elliott2011} or \citet{Pratt2001}. The P-R EOS is presented by Equation \ref{EQ: PR}.

	{\footnotesize
		\begin{equation}
			{\color{red}P}(t) = \cfrac{{\color{magenta}R}{\color{blue}T}(t,z)}{{\color{orange}v_m}({\color{blue}T}(t,z),{\color{red}P}(t)) - {\color{magenta}b}} - \cfrac{{\color{magenta}a^c}{\color{orange}\alpha}\left({\color{blue}T}(t,z)\right)}{{\color{orange}v_m}({\color{blue}T}(t,z),{\color{red}P}(t))^2 + 2{\color{magenta}b}{\color{orange}v_m}({\color{blue}T}(t,z),{\color{red}P}(t)) - {\color{magenta}b}^2}
			\label{EQ: PR}
		\end{equation}
	}

	The Equation \ref{EQ: PR} can be also written as the third-order polynomial equation shown in Equation \ref{eq: Compressibility}. Numerical methods such as Newton-Raphson can be used to solve the polynomial equation, but the Cardano formula is preferred as it yields an analytical solution. Further details regarding the Cardano formula can be found in Appendix \ref{CH: Cardano}.

@@

	{\footnotesize
	\begin{subequations} \label{eq: PR_AB}
		\begin{align} 
			{\color{orange}A}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]	& = \dfrac{{\color{orange}\alpha}\left[{\color{blue}T}\left(t,z\right)\right]{\color{magenta}a^c}{\color{red}P}(t)}{{\color{magenta}R}^2{\color{blue}T}^2(t,z)};													\label{eq: PR_A}\\
			{\color{orange}B}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]	& = \dfrac{{\color{magenta}b}{\color{red}P}(t)}{{\color{magenta}R}{\color{blue}T}(t,z)}	\label{eq: PR_B}.
		\end{align}
	\end{subequations}}

	The quantity ${\color{orange}\alpha}\left[{\color{blue}T}\left(t,z\right)\right]$ is a dimensionless correction term that depends on the temperature ${\color{blue}T}\left(t,z\right)$ and a constant ${\color{magenta}\kappa}$. The constant ${\color{magenta}\kappa}$ is given by the formula ${\color{magenta}\kappa} = 0.37464 + 1.54226 {\color{magenta}\omega} - 0.26992 {\color{magenta}\omega^2}$, where ${\color{magenta}\omega}$ is the acentric factor of $CO_2$ molecules and is equal to 0.239.

@@

	%	The compressibility factor can be obtained by solving the polynomial form of the P-R EOS given by Equation \ref{EQ: PR_polynomial}.

	%	{\footnotesize
		%		\begin{equation}
			%			Z^3 - (1-B)Z^2+(A-2B-3B^2)Z -(AB-B^2-B^3) = 0
			%			\label{EQ: PR_polynomial}
			%		\end{equation}
		%	}
	%	
	In a one-phase region, the fluid is described by one real root corresponding to the gas, liquid or supercritical phase. The gas-liquid mixture is present in the two-phase region, and two roots are found. The biggest root is assigned to the gas phase, and the smallest root corresponds to the liquid phase.

@@

	where ${\color{magenta}M_{CO_2}}$ denotes the molar mass of $CO_2$, and ${\color{orange}\overline{Z}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ is the compressibility factor that solves Eq. \eqref{eq: Compressibility}. This means that the density of the fluid depends on both space and time, due to its dependence on temperature and pressure.
%	, and it can be expressed in terms of a set of physical constants, denoted by $\varphi{\rho_\text{F}} = \left({\color{magenta}R}, {\color{magenta}M_{CO_2}}, {\color{magenta}T^c{CO_2}}, {\color{magenta}P^c{CO_2}}, {\color{magenta}\omega{CO_2}}\right)$.

@@

	The specific heat $C_p^{\text{F}}$ can be calculated from the equation of state, again under the assumption that the fluid phase consists of pure carbon dioxide and that the specific heat of real fluids can be calculated from an ideal contribution plus a residual term \citet{Pratt2001}. The main steps in the derivation of ${\color{orange}C^{CO_2}_p}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ are presented below:

	{\footnotesize
	\begin{subequations}\begin{align}
			{\color{orange}C^{CO_2}_v}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right] & = {\color{orange}C_v^{\text{I}}}\left[{\color{blue}T}(t,z),P(t)\right] + {\color{orange}C_v^{\text{R}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right] \label{eq: Cv_CO2}; \\ 
			{\color{orange}C^{CO_2}_p}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right] & = \underbrace{{\color{orange}C_p^{\text{I}}}\left[{\color{blue}T}(t,z),P(t)\right]}_{\text{Eq. } \eqref{eq: Cp_I}} + \underbrace{{\color{orange}C_p^{\text{R}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]}_{\text{Eq. } \eqref{eq: Cp_R}}  \label{eq: Cp_CO2}.
	\end{align}\end{subequations} }

	${\color{orange}C^{CO_2}_v}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ and ${\color{orange}C^{CO_2}_p}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ are the specific heat of $CO_2$ at constant volume and pressure, respectively. ${\color{orange}C_v^{\text{I}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ and ${\color{orange}C_p^{\text{I}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$, with ${\color{orange}C_p^{\text{I}}}({\color{blue}T}(t,z)) - {\color{orange}C_v^{\text{I}}}({\color{blue}T}(t,z)) = {\color{magenta}R}$, are the specific heat of an ideal gas at constant volume and pressure. ${\color{orange}C_v^{\text{R}}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ and ${\color{orange}C_p^{\text{R}}}({\color{blue}T}(t,z), {\color{red}P}(t))$ are the correction terms. \\

### 2023-06-20 SFE-article/Sections/introduction_imp.tex
26.05
20.06

	In recent years, the use of solvents for extracting natural substances from solid materials and liquids has been an area of significant interest for research and development. Supercritical fluids, which exhibit both gas- and liquid-like properties, have proven particularly useful in extraction processes due to their pressure-dependent dissolving power. Supercritical $CO_2$, in particular, is attractive because it is nontoxic, non-flammable, and non-corrosive. Furthermore, its critical point is relatively low compared to other fluids, making it a suitable alternative to traditional extraction techniques.

@@

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.95\columnwidth]{Figures/SFE_PFD_cropped.pdf}
		\caption{Process flow diagram ({\color{red}Font size}) }
		\label{fig: SFE_drawing}
	\end{figure}

@@

	The study is structured as follows: Chapter \ref{CH: Thermodynamic} provides a general discussion on supercritical fluids to familiarize the reader with their properties. Chapter \ref{CH:Governing_equations_chapter} introduces the general balance equations, which are derived under the assumption of Low-Mach number in Chapter \ref{CH:Low_Mach_chapter}. The simplified balance equations are combined with the extraction kinetic equation to develop the process model in Chapter \ref{CH: Extraction_model}. The maximum likelihood technique is presented in Chapter \ref{CH: Parameter_estimation} and is then combined with the process model. The dataset obtained from laboratory experiments is used to close the optimization problem in Chapter \ref{CH: Experiments}. Finally, Chapters \ref{CH: Results} and \ref{CH: Conclusion} discuss the results of the parameter estimations and simulation results.

>>>>

	The present study focuses on extracting essential oil from the caraway (Carum carvi L.) seeds with supercritical fluid extraction and the modelling of that process. Caraway is a biennial plant belonging to the Apiaceae family. It is widespread in Asia, Europe, and North Africa. The essential oil obtained from caraway finds its potential application as a fragrance ingredient in perfumes, liquors, and toothpaste. As presented by \citet{Hromis2015}, the dried caraway contains nearly~2.8–5\% essential oil, from which the main compounds are carvone, pinene, camphene, limonene, and carveol.

	The economic feasibility of the process is crucial when selecting appropriate technology. Conventional processes are frequently used for essential oil extraction, such as distillation and organic solvent extraction. The distillation process is carried out at a high temperature, which causes thermal degradation of thermolabile compounds. Considering that, alternatives techniques like supercritical fluid extraction gained popularity for extraction. Supercritical $CO_2$, in particular, is attractive due to its unique properties such as inflammable, non-toxic, low critical temperature and non-corrosive. Furthermore, its critical point is relatively low compared to other fluids, making it a suitable alternative to traditional extraction techniques. The supercritical fluids exhibit both gas- and liquid-like properties so pressure-depended dissolving power can be adjusted by operating conditions.

@@

	\begin{figure}[h!]
		\centering
		\includegraphics[trim = 5cm 12cm 8cm 6cm, clip,width=\columnwidth]{Figures/PFD.pdf}
		\caption{Process flow diagram}
		\label{fig: SFE_drawing}
	\end{figure}

@@

	The study is structured as follows: Chapter \ref{CH: Thermodynamic} provides a general discussion on supercritical fluids to familiarize the reader with their properties. Chapter \ref{CH:Governing_equations_chapter} introduces the general balance equations, which are derived under the assumption of Low-Mach number. The simplified balance equations are combined with the extraction kinetic equation to develop the process model in Chapter \ref{CH: Extraction_model}. The maximum likelihood technique is presented in Chapter \ref{CH: Parameter_estimation} and is then combined with the process model. The dataset obtained from laboratory experiments is used to close the optimization problem in Chapter \ref{CH: Experiments}. Finally, Chapters \ref{CH: Results} and \ref{CH: Conclusion} discuss the results of the parameter estimations and simulation results.
